---
title: "Plotting in R"
author: "Katie Cook"
date: "18 February 2019"
output:
  html_document:
    self_contained: no
---

#Tasks
1. Download Japanese protected areas from Protected Planet website (https://www.protectedplanet.net/) 
  (make sure you choose .shp file) and Unzip to your working directory folder.
   Download a Japanese map outline from an R package.  
2. Load spatial data into R.
3. Filter this data and make a beautiful map!
4. Have an explore and make some more maps! 

##Using the SF package 
The sf package is a new(ish) package for spatial data and analysis in R.
You can also use the sp package, but this doesn't work with dplyr and ggplot functions. 


## 1. Install and library packages, set your working directory to where your data is

```{r, include=TRUE}
knitr::opts_chunk$set(message = FALSE)
chooseCRANmirror(graphics=FALSE, ind=1)
#first install packages (if you haven't done this previously)

# devtools::install_github("tidyverse/ggplot2") #loads in the newest ggplot (for plotting)
 
# install.packages('dplyr') #for data management 

# install.packages('sf')  #'simple features' package for plotting and spatial data

# install.packages('ggspatial') #an add on to ggplot for spatial mapping

# install.packages('mapdata') #has data pre-installed for coastlines etc 

# install.packages('viridis') #nice color gradients


#now library the packages
library(ggplot2)
library(dplyr)
library(sf)
library(ggspatial)
library(mapdata)
library(viridis) 

setwd("//fbs4nas239/bskmc/Desktop Files/lab_meeting_map") # set working directory to where your files are

```

## 2. Loading in spatial data 

```{r, include=TRUE}

japanPA_point<- st_read('WDPA_Feb2019_JPN-shapefile-points.shp') #load in points 

japanPA_poly<-st_read('WDPA_Feb2019_JPN-shapefile-polygons.shp') #load in polygons

japan_coast<- st_as_sf(map('japan', plot=FALSE, fill=TRUE)) #get the japan map from mapdata

japan_sites<-read.csv('jp_waypoints.csv')


#Checking coordinate reference system is really important! Make sure you always do this!
st_crs(japanPA_point) #checks crs (coordinate reference system)
st_crs(japanPA_poly) 
st_crs(japan_coast)   

#all layers have crs EPSG 4326, which is WGS84 - unprojected data - ok. 
#if they are not the same you will have to use st_transform() to make them the same

```



## 3. Making spatial data and simple plots 

```{r}
#following on from last week (getting lat lon out of GPS) lets make this spatial 
japan_sites_sf<-st_as_sf(japan_sites, coords=c('lon', 'lat'), crs=4326)  #make sure the CRS is the same as your other data!

#now you can plot this as spatial data with attributes 
plot(japan_sites_sf)

#OR (without making it spatial)
ggplot(japan_sites, aes(x=lon, y=lat))+
  geom_point()





#now we can start plotting - basic plots to check 

plot(japanPA_point) #plots all the attributes
plot(japanPA_point$geometry) #plots just the geometry
plot (japanPA_point['IUCN_CAT'], key.pos = 1, key.width = lcm(1.3), key.length = 1.0, axes=TRUE) # colours points by designation, adds axes, adds key



#first lets make a basic map of japan using ggplot 
plot(japan_coast) # to check what it looks like

#plot this using ggplot
ggplot()+ geom_sf(data=japan_coast)+
  theme_bw()

#can fill map colours by ID 
ggplot()+ geom_sf(data=japan_coast, aes(fill=ID))+
  theme_bw()



```


##5. Using dplyr and ggplot

```{r}
#plotting all the polygons takes a long time- lets filter out some using dplyr
unique(japanPA_poly$MARINE) #gets the unique categories for the MARINE column

#1 is marine PAs, 0 is terrestrial PAs

#lets just plot marine PAs , need to filter the data - can use dplyr functions on sf objects
japan_marinePA<-filter(japanPA_poly, MARINE==1)  


#lets use ggplot to make it look nicer 
plot1<-ggplot(data=japan_marinePA)+
  geom_sf(color='cadetblue', fill='cadetblue')+         #plots the MPA shape, outline and fill color
  labs(title='Map of Japanese MPAs', y='Latitude', x='Longtitude')+ #adds labels
  theme_bw()+     #background theme
  annotation_scale(location = "tl")+   #scale bar (tl= top left) -part of ggspatial package
  annotation_north_arrow(location = "br", which_north = "true") #north arrow, br=bottom right


plot1   #calls the plots and plots it- might be quite slow.


#adding another layer- japan coast

#japan coastline map has prefecteral boundaries so remove these (like dissolve in arc map)
japan_coast<- japan_coast %>% st_buffer(0)  #add a buffer around all to make sure there are no gaps
japan_merge<-st_union(japan_coast) #merge it together, this creates a non-sf object
japan_merge<-st_sf(japan_merge)  #put it back to sf 
plot(japan_merge) #plot to check

#now we can combine the layers

#first create the base map
plot1<-ggplot(data=japan_merge)+    #the layer you want to plot first
  geom_sf( fill='antiquewhite',color='antiquewhite')+         # define outline and fill color
  labs(title='Map of Japanese MPAs', y='Latitude', x='Longtitude')+ #adds title and axis labels
  theme(panel.background = element_rect(fill = "lightblue",colour = "black",
                                        size = 0.5, linetype = "solid"), legend.title = element_blank())+     #background colour,legend theme
  annotation_scale(location = "tl")+   #scale bar (tl= top left) -part of ggspatial package
  annotation_north_arrow(location = "br", which_north = "true") #north arrow, br=bottom right

#now add the next layer (japanese MPA layer)
plot1<-plot1+geom_sf(data=japan_marinePA, aes( fill='cadetblue'), color='cadetblue')   #define colours
   #adds legend 

#finally add your study sites
plot1<-plot1+geom_sf(data=japan_sites_sf, aes(fill='red'), color='red')+
  scale_fill_identity( guide = 'legend', labels = c(' Japanese MPAs', 'Study Sites') ) 

#now plot
plot1


#plot maps with different CRS (UTM 54N)- appropriate projection for japan (Can find this info online)
ggplot(data = japan_merge) +
  geom_sf() +
  coord_sf(crs = '+proj=eqdc +lat_0=0 +lon_0=138 +lat_1=34 +lat_2=40 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs')+
  theme_bw()

#adding in names 

japan_prefectures<-st_centroid(japan_coast) #gets the centroids of the prefectures (where to put the labels)

japan_prefectures_coord<-st_coordinates(japan_prefectures) #get the co-ords

japan_prefectures<- cbind(japan_prefectures, japan_prefectures_coord) #bind the co-ords to the sf object

ggplot(data = japan_coast) +
  geom_sf(col='gray79', fill='antiquewhite') +
  geom_text(data= japan_prefectures,aes(x=X, y=Y, label=ID),
            color = "darkblue", fontface = "bold", check_overlap = FALSE, size=1.8) +
  annotate(geom = "text", x = 130 , y = 40, label = "Sea of Japan", 
           fontface = "italic", color = "grey22", size = 4)+
  theme_bw()+
  annotation_scale(location = "br")+  
  annotation_north_arrow(location = "tl", which_north = "true")+
  labs(x='Longitude', y='Latitude')
  

#plotting with heatmap 
#download this: https://simplemaps.com/data/jp-cities


jp_cities<-read.csv('jp.csv')

#get the  min and max population for plotting legend 
min<- jp_cities$population %>% na.omit() %>% min()
max<- jp_cities$population %>% na.omit() %>% max()  

#plot
ggplot(data = japan_coast) +
  geom_sf(col='gray79', fill='antiquewhite') +
  geom_point(data=jp_cities, aes(x=lng, y=lat, size=population, col=population))+
  scale_colour_viridis(option='plasma', limits=c(min, max), breaks=seq(min, max, by=5000000))+  
  scale_size_continuous(limits=c(min, max), breaks=seq(min, max, by=5000000))+
  guides(color= guide_legend(), size=guide_legend())+ #need to add this, and the same breaks for colour and size to combine legends
  annotation_scale(location = "tl")+   #scale bar (tl= top left) -part of ggspatial package
  annotation_north_arrow(location = "br", which_north = "true")+ #north arrow, br=bottom right
  theme_bw()          


#plotting maps based upon IUCN cat (multiple panels)
japan<-map_data('japan') #japan map for base with lat lon data

plot2<- ggplot()+
  geom_path(data=japan, aes(long, lat, group=group), size=0.25 )+  #geom path is the basemap
  geom_sf(data=japan_marinePA, col='cadetblue', fill='cadetblue')+
  facet_wrap(~IUCN_CAT)+
  theme_bw()

plot2

```

